---
title: "Palmetto Binary Logistic Regression"
author: "Sydney Rilum"
date: "2/3/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(broom)
library(stargazer)
library(kableExtra)

```

## Introduction
using the Florida palmetto data (palmetto.csv), use binary logistic regression to test feasibility of using variables plant height (height), canopy length (length), canopy width (width), and number of green leaves (green_lvs) to classify whether a palmetto is species Serenoa repens (species 1) or Sabal etonia (species 2).

## Data Exploration
```{r}
# Read in the data
palmetto <- read_csv(here("data", "palmetto.csv"),
                     col_types = cols(.default = 'c')) %>% # read in everything as character type to guarantee no parsing failures
  mutate(height = as.numeric(height)) %>% 
  mutate(width = as.numeric(width)) %>% 
  mutate(length = as.numeric(length)) %>% 
  mutate(green_lvs = as.numeric(green_lvs))

```

A section containing 2 - 3 finalized data visualizations (with figure captions) in which you explore differences in height, canopy length, canopy width, and green leaves for the two species. Add a figure caption to each. Below each data visualization, add a short (~1 sentence) statement with a takeaway from the visualization.

```{r, fig.align = "center"}
# Data visualizations


# add labs with units (cm)
# change to proper species name
ggplot(data = palmetto, aes(x = length, y = height)) +
  geom_point(aes(color = species)) 

ggplot(data = palmetto, aes(x = width, y = length)) +
  geom_point(aes(color = species)) 

ggplot(data = palmetto, aes(x = width, y = height)) +
  geom_point(aes(color = species)) 

ggplot(data = palmetto, aes(x = green_lvs, y = height)) +
  geom_point(aes(color = species)) +
  facet_wrap(~ species)

ggplot(data = palmetto, aes(x = green_lvs, y = length)) +
  geom_point(aes(color = species)) +
  facet_wrap(~ species)

ggplot(data = palmetto, aes(x = green_lvs, y = width)) +
  geom_point(aes(color = species),
              alpha = 0.5) +
  labs(x = "green leaves",
       y = "width (cm)",
       color = "Species")
  
ggplot(data = palmetto, aes(x = species, y = green_lvs)) +
  geom_jitter(aes(color = species))
```

## Binary Logistic Regression

```{r}
# Create a subset of data containing only variables needed for blr
palmetto_species <- palmetto %>% 
  select(species, height:green_lvs) %>% 
  mutate(species = as.factor(species)) 

# Check the levels
levels(palmetto_species$species)

# Binary Logistic Regression
palmetto_blr <- glm(species ~ height + length + width + green_lvs, 
                            data = palmetto_species, 
                            family = "binomial")

# Look at the model
palmetto_blr

summary(palmetto_blr)

# Get a tidy version w/ broom: 
palmetto_blr_tidy <- broom::tidy(palmetto_blr)

```

**Table 1.** 

<div align="center">
```{r, results = 'asis'}
# Present binary linear regression model outcome in a finalized regression table
stargazer(palmetto_blr, type = "html", 
          title = "Binary Linear Regression Results",
          dep.var.labels = "Species = Sabal etonia",
          covariate.labels = c("Height", "Length", "Width", "Green leaf count"))
```
</div>


## Evaluate Model Prediction Success

```{r}
# Create a new subset that converts log odds to a probability
palmetto_blr_fitted <- palmetto_blr %>% 
  broom::augment(type.predict = "response")

# Model visualization
# model_viz <- ggplot(data = palmetto_blr_fitted, aes(x = green_lvs, y = .fitted)) +
#   geom_point(aes(color = species)) +
#   geom_smooth(aes(color = species), se = FALSE) +
#   labs(x = "Green Leaves",
#        y = "Probability of Sabal etonia")

# Create a new subset with a column that lists which species the model would classify the plant as given a 50% probability cutoff (generated from the 4 predictor variables)
palmetto_blr_fitted_sub <- palmetto_blr_fitted %>% 
  select(species, .fitted) %>%
  mutate(predicted_species = case_when(  # make a new column for which species the model predicts using a 50% cutoff
    .fitted >= 0.5 ~ "2",
    .fitted < 0.5 ~ "1")) %>% 
  mutate(classification = case_when(  # make a new column that lists if model was correctly classified or not by model
    species == predicted_species ~ "correct",
    species != predicted_species ~ "incorrect")) %>% 
  mutate(species = as.numeric(species)) %>% 
  mutate(predicted_species = as.numeric(predicted_species)) %>% 
  group_by(species, classification) %>% 
  count(classification) 

# Make a table data subset
palmetto_table <- palmetto_blr_fitted_sub %>% 
  pivot_wider(names_from = classification,
              values_from = n) %>% 
  mutate(species = case_when(  # rename species 1 and 2 to real names
    species %in% 1 ~ "Serenoa repens",
    species %in% 2 ~ "Sabal etonia")) %>% 
  column_to_rownames(var = "species")

# Add another column to `palmetto_table` that displays percent correctly classified
palmetto_table['percent_correct'] = palmetto_table['correct'] / (palmetto_table['correct'] + palmetto_table['incorrect']) * 100
```


**Table 2.** 
```{r}
# Finalized table
kbl(palmetto_table, digits = 2,
      col.names = c("Correctly Classified", "Incorrectly Classified", "% Correctly Classified")) %>% 
  kable_styling(full_width = F, position = "center") %>% 
  column_spec(1, italic = TRUE)

```



**Data Citation:** [Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative.](https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5)
